---
title: "Fig.2D. Ventrobasal thalamus"
author: "Evgenii O. Tretiakov"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    toc: true
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  autodep        = TRUE,
  cache          = FALSE,
  cache.lazy     = FALSE,
  dev            = c("png", "pdf"),
  echo           = TRUE,
  error          = FALSE,
  fig.align      = "center",
  fig.width      = 14,
  fig.height     = 12,
  message        = FALSE,
  warning        = FALSE
)
Sys.setenv(RETICULATE_PYTHON = "/opt/python/3.8.8/bin/python")
# Load tidyverse infrastructure packages
suppressPackageStartupMessages({
  library(here)
  library(RColorBrewer)
  library(viridis)
  library(tidyverse)
  library(magrittr)
  library(stringr)
  library(skimr)
  library(future)
  library(zeallot)
  library(kableExtra)
  library(reticulate)
})
reticulate::use_condaenv("/opt/python/3.8.8/bin/python")

suppressPackageStartupMessages({
  library(Seurat)
  library(SeuratWrappers)
  library(SeuratDisk)
  library(sctransform)
  library(glmGamPoi)
  library(patchwork)
  library(UpSetR)
  library(ggstatsplot)
  library(qs)
  library(Scillus)
  library(scCustomize)
  library(Nebulosa)
  library(mrtree)
  library(gprofiler2)
})

# Set paths
src_dir    <- here('code')
data_dir   <- here('data')
output_dir <- here('output')
plots_dir  <- here(output_dir, 'figures')
tables_dir <- here(output_dir, 'tables')
source(here(src_dir, 'genes.R'))
source(here(src_dir, 'functions.R'))

# parallelisation
n_cores <- 16
plan(multisession, workers = n_cores)
options(future.globals.maxSize = 400000 * 1024^2) # 400Gb


# set seed
reseed <- 42
set.seed(seed = reseed)

# ggplot2 theme
theme_set(ggmin::theme_powerpoint())

```


Load dataset

```{r load}
srt <- LoadH5Seurat(here(data_dir, "THP7_clusters.h5Seurat"))
Idents(srt) <- "sep_clstr"
neurons <- subset(srt, subset = Rbfox3 > 0 | Elavl4 > 0 | Snap25 > 0 | Stmn2 > 0)
gneurons <- subset(neurons, subset = Gal > 0)
```

We check list of genes based on our prior knowledge.

```{r}
gene_of_interest <-
  c("Gal", "Slc17a7", "Slc6a4", "Ache", 
    "Prkaca", "Adcy1", "Grin1", 
    "Cck", "Oxt", "Npy", "Sst", "Avp")

genes_present <- Gene_Present(data = srt, gene_list = gene_of_interest)
genes_present
```

```{r pl-dotplot-all-rna, fig.align='center', fig.width=4, fig.asp = 0.618}
DotPlot_scCustom(
  seurat_object = srt,
  assay = "RNA",
  features = genes_present$found_features,
  flip_axes = TRUE,
  x_lab_rotate = TRUE,
  colors_use = viridis(n = 30, alpha = .75, direction = -1, option = "E"))
```

```{r pl-clusters, fig.align='center', fig.width=5, fig.asp = 0.618}
DimPlot_scCustom(
  srt,
  label = TRUE,
  repel = TRUE,
  pt.size = 2,
  figure_plot = TRUE
  ) + ggtitle("Clusters") + NoLegend()
```




Derive and filter matrix of neurons
```{r derive-neurons-mtxs}
mtx_neurons <-
    gneurons %>%
    GetAssayData("data", "RNA") %>%
    as.data.frame() %>%
    t()
rownames(mtx_neurons) <- colnames(gneurons)

# Filter features
filt_low_genes <-
    colSums(mtx_neurons) %>%
    .[. > quantile(., 0.4)] %>%
    names()
mtx_neurons %<>% .[, filt_low_genes]

min_filt_vector <-
    mtx_neurons %>%
    as_tibble() %>%
    select(all_of(filt_low_genes)) %>%
    summarise(across(.fns = ~ quantile(.x, .1))) %>%
    as.list %>%
    map(as.double) %>%
    simplify %>%
    .[colnames(mtx_neurons)]

# Prepare table of intersection sets analysis
content_mtx_neuro <-
    (mtx_neurons > min_filt_vector) %>%
    as_tibble() %>%
    mutate_all(as.numeric)
```

Visualise intersections sets that we are going to use (highlighted)
```{r plot-upset-neuro}
upset(as.data.frame(content_mtx_neuro),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Gad1", "Gad2", "Slc32a1", "Slc17a6"),
  queries = list(
    list(
      query = intersects,
      params = list("Gad1", "Gad2", "Slc32a1"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Slc17a6"),
      active = T
    )
  ),
  nintersects = 60,
  empty.intersections = "on"
)
```

```{r plot-upset-neuro-mod}
upset(as.data.frame(content_mtx_neuro),
  order.by = "freq",
  sets.x.label = "Number of cells",
  text.scale = c(2, 1.6, 2, 1.3, 2, 3),
  nsets = 15,
  sets = c("Gal", "Slc32a1", "Slc17a6"),
  queries = list(
    list(
      query = intersects,
      params = list("Gal", "Slc32a1"),
      active = T
    ),
    list(
      query = intersects,
      params = list("Gal", "Slc17a6"),
      active = T
    )
  ),
  nintersects = 60,
  empty.intersections = "on"
)
```


Make subset of stable neurons
```{r derive-subset-def-neurons}
gneurons$gaba_status <-
  content_mtx_neuro %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_all(.fns = ~ .x > 0)) %>%
  .$gaba

gneurons$gaba_expr <-
  content_mtx_neuro %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_any(.fns = ~ .x > 0)) %>%
  .$gaba

gneurons$glut_status <-
  content_mtx_neuro %>%
  select(Slc17a6) %>%
  mutate(glut = Slc17a6 > 0) %>%
  .$glut

neuro_fin <-
  subset(gneurons,
    cells = union(
      WhichCells(gneurons,
        expression = gaba_status == TRUE & glut_status == FALSE),
      WhichCells(gneurons,
        expression = glut_status == TRUE & gaba_expr == FALSE)))
```

## Check contingency tables for neurotransmitter signature
```{r check-contingency-neurotransmitter-test-neurons}
neuro_fin@meta.data %>%
  janitor::tabyl(glut_status, gaba_status)
```


Make splits of neurons by neurotransmitter signature 
```{r derive-subsets-by-neurotransmitter-neurons}
neuro_fin$status <- neuro_fin$gaba_status %>%
  if_else(true = "GABAergic",
    false = "glutamatergic")
Idents(neuro_fin) <- "status"
SaveH5Seurat(
  object    = neuro_fin,
  filename  = here(data_dir, "neuro_fin-THP7"),
  overwrite = TRUE,
  verbose   = TRUE
)

## Split on basis of neurotrans and test for difference
neuro_fin_neurotrans <- SplitObject(neuro_fin, split.by = "status")

```

## DotPlots grouped by orig.ident
Expression of GABA receptors in GABAergic Gal positive cells
```{r plot-dotplot-orig.ident-gabar-gaba-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$GABAergic,
        features = gabar,
        group.by = "orig.ident",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of GABA receptors in glutamatergic Gal positive cells
```{r plot-dotplot-orig.ident-gabar-glu-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$glutamatergic,
        features = gabar,
        group.by = "orig.ident",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in GABAergic Gal positive cells
```{r plot-dotplot-orig.ident-glur-gaba-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$GABAergic,
        features = glutr,
        group.by = "orig.ident",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in glutamatergic Gal positive cells
```{r plot-dotplot-orig.ident-glur-glu-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$glutamatergic,
        features = glutr,
        group.by = "orig.ident",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

## DotPlots 
Expression of GABA receptors in GABAergic Gal positive cells
```{r plot-dotplot-gabar-gaba-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$GABAergic,
        features = gabar,
        group.by = "k_tree",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of GABA receptors in glutamatergic Gal positive cells
```{r plot-dotplot-gabar-glu-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$glutamatergic,
        features = gabar,
        group.by = "k_tree",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in GABAergic Gal positive cells
```{r plot-dotplot-glur-gaba-neurons, fig.width=7, fig.asp=.5}
DotPlot(object = neuro_fin_neurotrans$GABAergic,
        features = glutr,
        group.by = "k_tree",
        cols = c("#adffff", "#0084ff"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```

Expression of glutamate receptors in glutamatergic Gal positive cells
```{r plot-dotplot-glur-glu-neurons, fig.width=6, fig.asp=.4}
DotPlot(object = neuro_fin_neurotrans$glutamatergic,
        features = glutr,
        group.by = "k_tree",
        cols = c("#ffc2c2", "#ff3c00"),
        col.min = -1, col.max = 1
) + RotatedAxis()
```


```{r drv-sbs-mtx-neurons}
sbs_mtx_neuro <-
    neurons %>%
    GetAssayData("data", "RNA") %>%
    as.data.frame() %>%
    t()
rownames(sbs_mtx_neuro) <- colnames(neurons)

# Filter features
filt_low_genes2 <-
    colSums(sbs_mtx_neuro) %>%
    .[. > quantile(., 0.4)] %>%
    names()
sbs_mtx_neuro %<>% .[, filt_low_genes2]

min_filt_vector2 <-
    sbs_mtx_neuro %>%
    as_tibble() %>%
    select(all_of(filt_low_genes2)) %>%
    summarise(across(.fns = ~ quantile(.x, .005))) %>%
    as.list %>%
    map(as.double) %>%
    simplify %>%
    .[filt_low_genes2]

# Prepare table of intersection sets analysis
content_sbs_mtx_neuro <-
    (sbs_mtx_neuro > min_filt_vector2) %>%
    as_tibble() %>%
    mutate_all(as.numeric)


neurons$gaba_status <-
  content_sbs_mtx_neuro %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_all(.fns = ~ .x > 0)) %>%
  .$gaba

neurons$gaba_expr <-
  content_sbs_mtx_neuro %>%
  select(Gad1, Gad2, Slc32a1) %>%
  mutate(gaba = if_any(.fns = ~ .x > 0)) %>%
  .$gaba

neurons$glut_status <-
  content_sbs_mtx_neuro %>%
  select(Slc17a6) %>%
  mutate(glut = Slc17a6 > 0) %>%
  .$glut

neuro_fin <-
  subset(neurons,
    cells = union(
      WhichCells(neurons,
        expression = gaba_status == TRUE & glut_status == FALSE),
      WhichCells(neurons,
        expression = glut_status == TRUE & gaba_expr == FALSE)))

neuro_fin$status <- neuro_fin$gaba_status %>%
  if_else(true = "GABAergic",
    false = "glutamatergic")
Idents(neuro_fin) <- "status"

sbs_mtx_neuro <-
    neuro_fin %>%
    GetAssayData("data", "RNA") %>%
    as.data.frame() %>%
    t()
rownames(sbs_mtx_neuro) <- colnames(neuro_fin)

# Filter features
filt_low_genes2 <-
    colSums(sbs_mtx_neuro) %>%
    .[. > quantile(., 0.4)] %>%
    names()
sbs_mtx_neuro %<>% .[, filt_low_genes2]

min_filt_vector2 <-
    sbs_mtx_neuro %>%
    as_tibble() %>%
    select(all_of(filt_low_genes2)) %>%
    summarise(across(.fns = ~ quantile(.x, .005))) %>%
    as.list %>%
    map(as.double) %>%
    simplify %>%
    .[filt_low_genes2]

# Prepare table of intersection sets analysis
content_sbs_mtx_neuro <-
    (sbs_mtx_neuro > min_filt_vector2) %>%
    as_tibble() %>%
    mutate_all(as.numeric)

```


```{r plot-piecharts-sbs-neuro-chisqr, fig.width=3, fig.asp=2.1}

sbs_mtx_neuro_full <- content_sbs_mtx_neuro |> 
  select(any_of(c(neurotrans, glutr, gabar, "Gal", "Galr1"))) |> 
  dplyr::bind_cols(neuro_fin@meta.data)

sbs_mtx_neuro_full |> glimpse()

sbs_mtx_neuro_full$Gal_sign <- 
  sbs_mtx_neuro_full %>%
  select(Gal) %>%
  mutate(Gal_sign = if_any(.fns = ~ .x > 0)) %>%
  .$Gal_sign

# for reproducibility
set.seed(reseed)
# plot
grouped_ggpiestats(
  # arguments relevant for `ggpiestats()`
  data = sbs_mtx_neuro_full,
  x = Gal_sign,
  grouping.var = status,
  perc.k = 1,
  package = "ggsci",
  palette = "category10_d3",
  # arguments relevant for `combine_plots()`
  title.text = "Molecular specification of thalamic neuronal lineages by Gal signalling and main neurotransmitter expression",
  caption.text = "Asterisks denote results from proportion tests; \n***: p < 0.001, ns: non-significant",
  plotgrid.args = list(nrow = 2)
)
```



```{r pl-feature-slc17a6, fig.align='center', fig.width=5, fig.asp = 0.618}
p1 <- FeaturePlot_scCustom(
  srt, "Slc17a6",
  pt.size = 2,
  order = TRUE,
  alpha_exp = 0.65,
  alpha_na_exp = 0.2,
  label = TRUE,
  repel = TRUE,
  colors_use = srt@misc$expr_Colour_Pal) + 
  ggtitle("Slc17a6(Vglut2): ") + theme(plot.title = element_text(size = 24))
p1
```

```{r pl-density-slc17a6, fig.align='center', fig.width=7, fig.asp = 0.618}
Plot_Density_Custom(
  seurat_object = srt,
  features = "Slc17a6",
  pt.size = 2) + 
  ggtitle("Slc17a6(Vglut2): ") + theme(plot.title = element_text(size = 24))
```

```{r pl-density-slc17a6-gal, fig.align='center', fig.width=7, fig.asp = 0.618}
Plot_Density_Joint_Only(
  seurat_object = srt,
  features = c("Slc17a6", "Gal"),
  pt.size = 2) + 
  ggtitle("Slc17a6(Vglut2) + Gal ") + theme(plot.title = element_text(size = 24))
```
```{r}
srt_vglut2 <- subset(srt, idents = c(2, 4, 5, 6, 7), invert = TRUE)
srt_vglut2 <- subset(srt_vglut2, subset = UMAP_1 > -15 & UMAP_1 < 2)
```

```{r pl-feature-gal, fig.align='center', fig.width=8, fig.asp = 0.618}
p2 <- FeaturePlot_scCustom(
  srt_vglut2, "Gal",
  pt.size = 5,
  order = TRUE,
  alpha_exp = 0.75,
  alpha_na_exp = 0.05,
  colors_use = srt@misc$expr_Colour_Pal) + 
  ggtitle("Gal in Vglut2+ populations") + theme(plot.title = element_text(size = 24))
p2
```

```{r pl-density-gal, fig.align='center', fig.width=7, fig.asp = 0.618}
Plot_Density_Custom(
  seurat_object = srt_vglut2,
  features = "Gal",
  pt.size = 2) + 
  ggtitle("Gal across Vglut cell-populations only") + theme(plot.title = element_text(size = 24))
```

```{r pl-density-gal-across-vglut2, fig.align='center', fig.width=7, fig.asp = 0.618}
Plot_Density_Joint_Only(
  seurat_object = srt_vglut2,
  features = c("Slc17a6", "Gal"),
  pt.size = 2) + 
  ggtitle("Slc17a6 + Gal across Vglut cell-populations only") + theme(plot.title = element_text(size = 24))
```

```{r pl-clusters-glut2, fig.align='center', fig.width=7, fig.asp = 0.618}
DimPlot_scCustom(
  srt_vglut2,
  label = TRUE,
  repel = TRUE,
  pt.size = 3,
  figure_plot = TRUE
  ) + ggtitle("Vglut2+ clusters") + NoLegend()
```

```{r pl-heatmap-vglut, fig.align='center', fig.width=8, fig.asp = 0.618}
DoHeatmap(srt_vglut2, features = genes_present$found_features)
```

```{r pl-dotplot-vglut2-rna, fig.align='center', fig.width=3, fig.asp = 0.618}
p3 <- DotPlot_scCustom(
  seurat_object = srt_vglut2,
  assay = "RNA",
  features = genes_present$found_features,
  flip_axes = TRUE,
  x_lab_rotate = TRUE,
  colors_use = viridis(n = 30, alpha = .75, direction = -1, option = "E"))
p3
```

```{r pl-fig-2d, fig.align='center', fig.width=18, fig.asp = 0.206}
p1 | p2 | p3
```
